---
title: "Cod Models"
author: "Becca Van Hoeck"
date: "4/19/2021"
output: html_document
---

Research Questions: 

1. Is the timing of Atlantic cod spawning-associated vocalization, including grunt presence and rate, correlated with environmental variables? 
2. Do these environmental correlates differ between the Massachusetts Bay and Cox Ledge Region? 

Data Source:
- Fixed station passive acoustic monitoring data from spawning season of 2013 and 2014 at Cox Ledge and Mass Bay
- Cox Ledge dataset consists of 1 site, whereas Mass Bay has 5 sites within the region. 
- Data are summarized by cod grunt presence and grunt rate per hour during the spawning season.

Current Method: 

- treat natural cycles as circular variables to estimate magnitude and direction of correlation
- Use GLMs to model grunt presence and rate separately
- Ultimately, I planned to use a hurdle model, but was having difficulty interpreting the output so I moved forward with separate models. 

Questions: 

1. Is this statistical framework suitable to evaluate the question? 
2. Am I building and evaluating the models correctly? Specifically, I am unsure how to account for the nested site within region and the unbalanced number of sites between the regions. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lunar)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

```

## Script to build models and evaluate fit 

Load data - all data wrangling and exploratory visualizations were done in historical_cod_data.RMD
```{r}

## Tidy data with grunt presence, rate, and associated environmental variables
# note that there are many times missing temperatures. Possibly not able to use this variable because grunt data is so sparse. I don't want to remove some cod data just to have complete cases with temperature

cod_data = read.csv("data/cod_allregions07-21.csv", header = TRUE)
cod_data = cod_data[,-1] # remove row index

```

## Creating sin and cos variables for natural cycles

```{r}

# Hour of the day
cod_data$shour = sin(2*pi*(cod_data$hour/24)) 
cod_data$chour = cos(2*pi*(cod_data$hour/24)) 

# Julian Day
cod_data$sday = sin(2*pi*(cod_data$J/365))
cod_data$cday = cos(2*pi*(cod_data$J/365))
  
# Lunar cycle  
cod_data$sLunar = sin(cod_data$lunarphase) 
cod_data$cLunar = cos(cod_data$lunarphase)

#Semi lunar cycle
cod_data$sLunar2 = sin(cod_data$lunar2)
cod_data$cLunar2 = cos(cod_data$lunar2)


```

# Modeling presence and rate separately

# Presence with Cox Ledge and Mass Bay Data

Full presence mod from caiger et al: Y + S + H + D + L1 + L2 +J + rW + S:J
(Caiger et al. evaluated timing of cod grunts with a longer time-series dataset in Mass Bay)

```{r}
library(lme4)

# singularity for region when site is included
codMod = glm(presence ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + sday + cday,
                 data = cod_data, family = binomial(link = "logit"))
summary(codMod)
drop1(codMod)

# remove site, but add region
codMod_nosite = glm(presence ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + sday + cday,
                 data = cod_data, family = binomial(link = "logit"))
summary(codMod_nosite)
drop1(codMod_nosite)

# Same as CodMod_nosite, but added site as a random intercept
codMod_rSite = glmer(presence ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + sday + cday + 
                       (1|site), data = cod_data, family = binomial(link = "logit"))
summary(codMod_rSite)
# region is highly correlated with the intercept and its fixed effect  p value is high
# i don't know how to interpret the random effects

# Random site nested within Region - not sure if code is correct. The summary is the same as rSite model above
codModNB_rSiteNested = glmer(presence ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + 
                                sday + cday + (1|site:region), data = cod_data, family = binomial(link = "logit"))
summary(codModNB_rSiteNested)

# Random slope and intercept for site? again not sure code is correct
# warning: boundary (singular) fit
codModNB_rSiteIS = glmer(presence ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + 
                                sday + cday + (1+site|region), data = cod_data, family = binomial(link = "logit"))
summary(codModNB_rSiteIS)# change to cod mod once run

# AIC suggests that the simplest model - without region is the best fit
AIC(codMod, codMod_nosite, codMod_rSite, codModNB_rSiteNested, codModNB_rSiteIS)


```

## Model Grunt Rate

Rate mod from Caiger et al: Y, S, H, L1, L2, J, rW, S:J.
```{r}
#library(gamlss)
library(lme4)
library(MASS) # has glmer.nb function

# Full model - no region 
RateMod_full = glm(n_grunts ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2, 
               data = cod_data, poisson(link = "log"))
summary(RateMod_full)

# same model but negative binomial 
RateModNB_full = glm.nb(n_grunts ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2,
                   data = cod_data)
summary(RateModNB_full)

RateModNB_region = glm.nb(n_grunts ~ SpawnSeason + region + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2,
                   data = cod_data)
summary(RateModNB_region)

# AIC suggests the neg bin is a much better fit
AIC(RateMod_full, RateModNB_full)

# Continue using NB error, add region and change site to a random intercept
RateModNB_rSite = glmer.nb(n_grunts ~ SglmpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 +
                       (1|site), data = cod_data)
summary(RateModNB_rSite)
# Region was not significant
# no strong correlations between predictor variables - suggests data do not need to be centered or rescaled
# p value given in fixed effects table, but not usually given for lmer output. Can I trust those values?

drop1(RateModNB_rSite, test = "Chisq")

library(lmerTest)
drop1(RateModNB_rSite, ddf = "Satterthwaite")

# mixed effects model has higher AIC than neg bin with just site
AIC(RateMod_full, RateModNB_full,RateModNB_rSite)

RateModNB_rSiteNested = glmer.nb(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 +
                       (1|site:region), data = cod_data)
summary(RateModNB_rSiteNested)

RateModNB_rSiteIS = glmer.nb(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 +
                       (1+site|region), data = cod_data)
summary(RateModNB_rSiteIS)
 
#ratePredict = predict(RateModNBII)
#plot(ratePredict)
```

# Mixture and hurdle models
Caiger et al 2020 (MEPS) Global Model: Ten years of data from Mass Bay

presence or rate ~ year + site + hour + depth + lunar + semiLunar + JulianDay + S:Julian Day + random(week)

- hour, lunar, semilunar, julian day were all cyclical
- random effect of week was to account for serial autocorrelation
- Used model averaged predictions, weighted by AIC, with all models with deltaAIC <5. 

Presence model: Top 4 models averaged. All included Y, S, H, D, J, rW, S:J. One included L1, one omitted all Lunar, one included both L1 & L2, one included only L1. 

Rate model: model evaluated predictors for the rate zero-inf terms separately. Two models were averaged, rate terms were the same and included: Y, S, H, L1, L2, J, rW, S:J. Zero inflated were the same as rate, except one model excluded L2. 

```{r}
# following ENEC563 Lab 15 for Mixture and Hurdle Models

# starting with the global model
# does not include region yet, because of singularity with site
# also does not include random effect of week. 

# library(pscl)
# zip1 = zeroinfl(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sday + cday + site:J |
#                   SpawnSeason + region + shour + chour + sLunar + cLunar + sday + cday + site:J, 
#                 dist = "poisson", link = "logit", data = cod_data)
# summary(zip1)
# 
# #this was computationally singular? 
# Nb1 = zeroinfl(n_grunts ~ SpawnSeason + region + shour + chour + sLunar + cLunar + sday + cday + site:J |
#                   SpawnSeason + region + shour + chour + sLunar + cLunar + sday + cday + site:J, 
#                 dist = "negbin", link = "logit", data = cod_data)
# summary(Nb1)
# 
# # but the likelihood is much higher for the Neg Bin that only gave coefficient estimates
# library(lmtest)
# lrtest(zip1, Nb1)


```


