---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Misc cod model code



```{r}

# Calculating and comparing residuals

# this code came from class notes, but don't remember which lesson
cod_data$P_resid = residuals(codMod, type = "pearson")
cod_data$d_resid = residuals(codMod, type = "deviance")

sum(cod_data$P_resid)/df.residual(codMod_nosite)
# does this mean that there is a significant difference in presence, or significant difference in covariates?


```

## Model each region separately and compare best models

Because I am not sure the above method is appropriate for evaluating whether the environmental correlates differ between the two regions, I modeled each region separately. 

This workflow leads to the same final model for each region. This adds some confirmation to me that there likely is no significant difference between the timing of spawning in the two regions. 

```{r}
CoxLedge = cod_data %>%
  filter(region == 'CoxLedge')

MassBay = cod_data %>%
  filter(region == "MassBay")

# Cox ledge data

# Presence
codModCL = glm(presence ~ SpawnSeason + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + sday + cday,
                 data = CoxLedge, family = binomial(link = "logit"))
summary(codModCL)
drop1(codModCL)

# Rate
# Full model - no region 
RateMod_fullCL = glm(n_grunts ~ SpawnSeason + shour + chour + sLunar + cLunar + sLunar2 + cLunar2, 
               data = CoxLedge, poisson(link = "log"))
summary(RateMod_fullCL)

# same model but negative binomial 
library(MASS)
RateModNB_fullCL = glm.nb(n_grunts ~ SpawnSeason + shour + chour + sLunar + cLunar + sLunar2 + cLunar2,
                   data = CoxLedge)
summary(RateModNB_fullCL)

# AIC suggests the neg bin is a much better fit
AIC(RateMod_fullCL, RateModNB_fullCL)

drop1(RateModNB_fullCL)



```

# Mass Bay data

```{r}

# Presence
codModMB = glm(presence ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2 + sday + cday,
                 data = MassBay, family = binomial(link = "logit"))
summary(codModMB)

drop1(codModMB)

# Rate
# Full model - no region 
RateMod_fullMB = glm(n_grunts ~ SpawnSeason + site + shour + chour + sLunar + cLunar + sLunar2 + cLunar2, 
               data = MassBay, poisson(link = "log"))
summary(RateMod_fullMB)

# same model but negative binomial 
library(MASS)
RateModNB_fullMB = glm.nb(n_grunts ~ SpawnSeason + shour + chour + sLunar + cLunar + sLunar2 + cLunar2,
                   data = MassBay)
summary(RateModNB_fullMB)

# AIC suggests the neg bin is a much better fit
AIC(RateMod_fullMB, RateModNB_fullMB)

drop1(RateModNB_fullMB)



```
